---
title: "ENVS-193DS_homework-03"
author: "Shunji Shimoyoshi"
format: pdf
execute:
  message: false # use this to make sure messages don't show up
  warning: false # use this to make sure warnings don't show up
---

# Homework Setup

```{r}
#| label: reading in packages and data

# reading in packages
library(tidyverse)
library(here)
library(janitor)
library(readxl)
# creating new object from data set
kelp <- read_csv(here("data","temp-kelp.csv"))
bike <- read_csv(here("data","ENVS-193DS-Personal-Data-Project-cont.csv"))

```

# Problem 1. Giant kelp fronds

## a. An appropriate test

The two tests that are appropriate to describe the strength of the relationship between temperature and giant kelp frond elongation rate are a linear model and a correlation test. A linear model will tell us how well temperature can be used to predict kelp frond elongation rate, while correlation can tell us whether these two variables are related to each other. Given that we do not have a specific hypothesis on how temperature affects kelp frond elongation rate, both tests should be used to accurately describe the relationship. 

## b. Create a visualization

```{r}
#| label: visualizing data

# base layer: ggplot
ggplot(data = kelp,
       # x-axis: temperature
       # y-axis: kelp frond elongation rate
       mapping = aes(x = temp_c,
                     y = kelp_elong)) +
  # first layer: points to represent individual observations
  geom_point(color = "blue") + # using custom colors
  # relabeling axes, adding title
  labs(x = "Temperature (°C)", 
       y = expression("Kelp frond elongation rate(cm day"^-1~")")) +
  # changing theme
  theme_light()

```

## c. Check your assumptions and run your test

### Checking assumptions

```{r}
#| label: kelp qq plot

# base layer: ggplot
ggplot(data = kelp, # data frame: kelp
       # y-axis for QQ plot (no x-axis for QQ plot)
       aes(sample = kelp_elong)) + 
  # first layer: QQ reference line
  geom_qq_line(color = "blue") +
  # second layer: QQ plot
  geom_qq() +
  # custom theme
  theme_light()

```

Based on the initial visualization, we can expect a linear relationship between temperature and kelp frond elongation rate. Both temperature and kelp frond elongation rate are continuous variables. Both variables are normally distributed, based on points roughly following the reference line on the QQ plot. Finally, observations are assumed to be independent.

### Running test

```{r}
#| label: correlation test for kelp model

cor.test(kelp$kelp_elong, kelp$temp_c, # variables: kelp elongation rate and temperature
         method = "pearson") # method: pearson's correlation (r)

```

## d. Results communication

To evaluate the strength of the relationship between temperature and giant kelp frond elongation rate, I used a Pearson's correlation (r) test. A correlation test was appropriate for this as I wanted to determine the strength of the relationship between these two variables. 

I found a moderate, negative relationship between ocean temperature and giant kelp frond elongation rate (Pearson's r = -0.69, t(30) = -5.12, p < 0.001, $\alpha$ = 0.05).










```{r}
#| label: checking assumptions of kelp model

#par(mfrow = c(2, 2))
#plot(kelp_model)

```

- Residuals vs Fitted and Scale-Location: checking homoscedasticity of residuals, residuals seem to have constant variance based on even distribution around the horizontal red line

- Q-Q Residuals: checking distribution of residuals, residual is normally distributed based on points following the dotted line relatively closely

- Residuals vs Leverage: checking for outliers, no outliers outside dotted lines that would influence model predictions







### Running test

```{r}
#| label: kelp model

#kelp_model <- lm(
#  kelp_elong ~ temp_c, # formula: kelp elongation rate as a function of temperature
#  data = kelp # data frame: kelp
#)

```

### Model Summary

```{r}
#| label: summary of kelp model

# displaying result of kelp linear model
#summary(kelp_model)

```

## d. Results communication

To evaluate 




## e. Test implications

## f. Double check your work






