---
title: "ENVS 193DS Homework 3"
author: "Shunji Shimoyoshi"
format: pdf
execute:
  message: false # use this to make sure messages don't show up
  warning: false # use this to make sure warnings don't show up
---

# Homework Setup

```{r}
#| label: reading in packages and data

# reading in packages
library(tidyverse)
library(here)
library(janitor)
library(readxl)
# creating new object from data set
kelp <- read_csv(here("data","temp-kelp.csv"))
bike <- read_csv(here("data","ENVS-193DS-Personal-Data-Project-cont.csv"))

```

# Problem 1. Giant kelp fronds

## a. An appropriate test

The two tests that are appropriate to describe the strength of the relationship between temperature and giant kelp frond elongation rate are a linear model and a correlation test. A linear model will tell us how well temperature can be used to predict kelp frond elongation rate, while correlation can tell us whether these two variables are related to each other. Given that we do not have a specific hypothesis on how temperature affects kelp frond elongation rate, both tests should be used to accurately describe the relationship. 

## b. Create a visualization

```{r}
#| label: visualizing data

# base layer: ggplot
ggplot(data = kelp, # data frame: kelp
       # x-axis: temperature
       # y-axis: kelp frond elongation rate
       mapping = aes(x = temp_c,
                     y = kelp_elong)) +
  # first layer: points to represent individual observations
  geom_point(color = "blue") + # using custom colors
  # relabeling axes, adding title
  labs(x = "Temperature (°C)", 
       y = expression("Kelp frond elongation rate(cm day"^-1~")")) +
  # changing theme
  theme_light()

```

## c. Check your assumptions and run your test

### Checking assumptions

```{r}
#| label: kelp qq plot

# base layer: ggplot
ggplot(data = kelp, # data frame: kelp
       # y-axis for QQ plot (no x-axis for QQ plot)
       aes(sample = kelp_elong)) + 
  # first layer: QQ reference line
  geom_qq_line(color = "blue") +
  # second layer: QQ plot
  geom_qq() +
  # custom theme
  theme_light()

```

Based on the initial visualization, we can expect a linear relationship between temperature and kelp frond elongation rate. Both temperature and kelp frond elongation rate are continuous variables. Both variables are normally distributed, based on points roughly following the reference line on the QQ plot. Finally, observations are assumed to be independent.

### Running test

```{r}
#| label: correlation test for kelp model

cor.test(kelp$kelp_elong, kelp$temp_c, # variables: kelp elongation rate and temperature
         method = "pearson") # method: pearson's correlation (r)

```

## d. Results communication

To evaluate the strength of the relationship between temperature and giant kelp frond elongation rate, we used a Pearson's correlation (r) test. A correlation test was appropriate for this as we wanted to determine the strength of the relationship between these two variables. 

We found a moderate negative relationship between ocean temperature and giant kelp frond elongation rate (Pearson's r = -0.69, t(30) = -5.12, p < 0.001, $\alpha$ = 0.05).

## e. Test implications

From this study, we found that as ocean temperature increases, giant kelp frond elongation rate decreases. We found a moderate negative relationship between ocean temperature and giant kelp frond elongation rate (Pearson's r = -0.69, t(30) = -5.12, p < 0.001, $\alpha$ = 0.05).

## f. Double check your work

### Fitting linear model

```{r}
#| label: kelp linear model

kelp_model <- lm(
  kelp_elong ~ temp_c, # formula: kelp elongation rate as a function of temperature
  data = kelp # data frame: kelp
)

```

### Diagnostics

```{r}
#| label: checking assumptions of kelp model

# display plots in a 2x2 grid
par(mfrow = c(2, 2))
# display diagnostic plots
plot(kelp_model)

```

- Residuals vs Fitted and Scale-Location: checking homoscedasticity of residuals, residuals seem to have constant variance based on even distribution around the horizontal red line

- Q-Q Residuals: checking distribution of residuals, residual is normally distributed based on points following the dotted line relatively closely

- Residuals vs Leverage: checking for outliers, no outliers outside dotted lines that would influence model predictions

### Model Summary

```{r}
#| label: summary of kelp model

# displaying result of kelp linear model
summary(kelp_model)

```

**For each 1°C of ocean temperature increase, we would expect a giant kelp frond elongation rate decrease of 0.43 ± 0.08 (SE) cm/day.**

This result from the linear model would have led me to make the same decision about the null hypothesis. As ocean temperature increases, giant kelp frond elongation rate decreases. The only difference is that the strength of this relationship would have not been known without the correlation test.

# Problem 2. Personal Data

## a. Updating your visualizations

```{r}
#| label: visualizing personal data project

# base layer: ggplot
ggplot(data = bike, # data frame: bike
       # x-axis: weather
       # y-axis: Duration of bike ride
       # coloring by weather
       aes(x = Weather,
           y = `Duration of bike ride (minutes)`,
       color = Weather)) +
  # first layer: boxplot
  geom_boxplot() +
 # relabeling the axes, adding title and subtitle
  labs(x = "Wind Condition", 
       y = "Duration of bike ride (minutes)",
       title = "Biking in windy conditions increases the time spend biking",
       subtitle = "Most recent observation: 5/23/2026") +
 # custom theme
  theme_light() +
  # removing legend
  theme(legend.position = "none") + 
  # changing the color
  scale_color_manual(values = c("calm" = "blue",
                                "windy" = "red"))

```

```{r}
#| label: visualizing personal data project 2

# base layer: ggplot
ggplot(data = bike, # data frame: bike
       # x-axis: duration of day's activities
       # y-axis: Duration of bike ride
       aes(x = `Duration of day's activities (hours)`,
           y = `Duration of bike ride (minutes)`,)) +
  # first layer: points
  geom_point(color = "firebrick") +
 # relabeling the axes, adding title and subtitle
  labs(x = "Duration of day's activities (hours)", 
       y = "Duration of bike ride (minutes)",
       title = "Daily Duration of Activities Effect on Time Spent Biking",
       subtitle = "Most recent observation: 5/23/2026") +
 # custom theme
  theme_light() +
  # removing legend
  theme(legend.position = "none")

```

## b. Captions

**Figure 1. Biking in windy conditions increases the time spend biking** Difference in time spent biking between calm and windy conditions is shown as a boxplot. Median, IQR, errorbars, and outliers are visualized and wind condition type is differentiated based on color. Wind data citation: Weather Underground SBFD Station 1 - KCASANTA4905

**Figure 2. Daily Duration of Activities Effect on Time Spent Biking** As duration of day's activities increases, time spent biking also increases. Individual, daily observations are shown as dots on the graph.



